Active Transfer Learning

### 2018-03-15 paper/paper.tex
Edits to the writeup

We define the Bellman error (or Bellman residual) of a $Q$-function $Q$ as $B(Q) \triangleq TQ - Q$. Notice that a $Q$-function $Q$ is optimal if, and only if, $||B(Q)||_{\infty} = 0$.

@@

We assume states and actions to be drawn from a fixed joint distribution $\mu$. We define the set of $Q$-functions of our interest as the set $\mathcal{Q}^{\epsilon}$ of all $Q$ functions whose Bellman error $B(Q)$ is in some $\epsilon$-ball defined by the $l_p$-norm $|| \cdot ||_{p,\mu}$:
\begin{equation}
\mathcal{Q}^{\epsilon} = \left\{ Q\in\mathcal{Q} \mid || B(Q) ||_{p,\mu}^p \leq \epsilon \right\}
\end{equation}
Given a dataset of $N$ samples $D = \langle s_i,a_i,r_i,s_i' \rangle_{i=1}^N$, we can approximate the $l_p$-norm of the Bellman error of a $Q$-function $Q$ as:
\begin{equation}
||B(Q)||_{p,D}^p = \frac{1}{N}\sum_{i=1}^N \left\lvert r_i + \gamma\max_{a'}Q(s_i',a') - Q(s_i,a_i) \right\rvert^p
\end{equation}

\begin{theorem}
Let $Q$ be a $Q$-function with empirical Bellman error, computed on a dataset $D$ of $N$ i.i.d. samples, given by $||B(Q)||_{p,D}^p = \hat{q}$. Then, for any $\epsilon >= 0$:
\begin{equation}
P\left(Q\in\mathcal{Q}^{\epsilon} \mid ||B(Q)||_{p,D}^p = \hat{q}\right) \leq exp\left( -\frac{2N\max\{\epsilon-\hat{q},0\}^2}{\left(\frac{2R_{max}}{1-\gamma}\right)^{2p}} \right)
\end{equation}
\end{theorem}
\begin{proof}
Assume $\hat{q}>\epsilon$. Then:
\begin{align*}
P\left(Q\in\mathcal{Q}^{\epsilon} \mid ||B(Q)||_{p,D}^p = \hat{q}\right) =\ & 
P\left(|| B(Q) ||_{p,\mu}^p \leq \epsilon \mid ||B(Q)||_{p,D}^p = \hat{q}\right) \\ =\ &
P\left(|| B(Q) ||_{p,\mu}^p - \hat{q} \leq \epsilon - \hat{q} \mid ||B(Q)||_{p,D}^p = \hat{q}\right)
\end{align*}
Notice that $\mathbb{E}[\hat{q}] = || B(Q) ||_{p,\mu}^p$ and that $\epsilon - \hat{q} < 0$ by assumption. Then, we can apply Hoeffding's inequality to write:
\begin{equation*}
P\left(Q\in\mathcal{Q}^{\epsilon} \mid ||B(Q)||_{p,D}^p = \hat{q}\right) \leq exp\left( -\frac{2N(\epsilon-\hat{q})^2}{\left(\frac{2R_{max}}{1-\gamma}\right)^{2p}} \right)
\end{equation*}
Finally, when $\hat{q} \leq \epsilon$, the probability can be straightforwardly upper-bounded by $1$. Combining the two results concludes the proof.
\end{proof}
\begin{theorem}
Let $Q$ be a $Q$-function with empirical Bellman error, computed on a dataset $D$ of $N$ i.i.d. samples, given by $||B(Q)||_{p,D}^p = \hat{q}$. Then, for any $\epsilon >= 0$:
\begin{equation*}
P\left( ||B(Q)||_{p,D}^p = \hat{q} \mid Q\in\mathcal{Q}^{\epsilon} \right) \leq \frac{\epsilon}{\hat{q}}
\end{equation*}
\begin{proof}
\begin{align*}
P\left( ||B(Q)||_{p,D}^p = \hat{q} \mid Q\in\mathcal{Q}^{\epsilon} \right) =\ &
P\left( ||B(Q)||_{p,D}^p = \hat{q} \mid || B(Q) ||_{p,\mu}^p \leq \epsilon \right)\\ \leq\ &
P\left( ||B(Q)||_{p,D}^p \geq \hat{q} \mid || B(Q) ||_{p,\mu}^p \leq \epsilon \right)\\ \leq\ &
\frac{E[||B(Q)||_{p,D}^p]}{\hat{q}}\\ \leq\ &
\frac{\epsilon}{\hat{q}}
\end{align*}
The first inequality is straightforward, the second one is from Markov's inequality, while the third one is due to the fact that $Q\in\mathcal{Q}^{\epsilon}$.
\end{proof}
\end{theorem}

@@

Consider solving the target task given a dataset $D$ of $N$ samples. The posterior distribution over optimal $Q$-functions is:
\begin{equation}
P(Q^* \mid D) \propto P(D \mid Q^*)P(Q^*)
\end{equation}
Bounding the likelihood $P(D \mid Q^*)$ ({\color{red}FIND A BETTER DERIVATION}) and taking the maximum of the log-posterior, we obtain the following optimization problem:
\begin{equation}
\min_{Q\in\mathcal{Q}} ||B(Q)||_{p,D}^p - \log P(Q) 
\end{equation}
Then, let us specify a particular hypothesis space $\mathcal{Q}$. We consider $Q$-functions $Q_{\bm{w}}$ parameterized by the vector $\bm{w}$. Our optimization problem becomes:
\begin{equation}
\min_{\bm{w}} ||B(\bm{w})||_{p,D}^p - \log P(\bm{w}) 
\end{equation}
We model the prior distribution over the optimal parameters $\bm{w}$ as a Gaussian. That is, we assume:
\begin{equation}
P(\bm{w}) = \mathcal{N}(\bm{\mu},\bm{\Sigma})
\end{equation}
Then, our optimization, adopting the $l_2$-norm, becomes:
\begin{equation} \label{eq:optim}
\min_{\bm{w}} ||B(\bm{w})||_{D}^2 + ||\bm{w}-\bm{\mu}||_{\bm{\Sigma}} = \min_{\bm{w}} \frac{1}{N}\sum_{i=1}^N\left\lvert y_i - Q_{\bm{w}}(s_i,a_i) \right\rvert^2 + ||\bm{w}-\bm{\mu}||_{\bm{\Sigma}}
\end{equation}
where $y_i = r_i + \gamma\max_{a'}Q_{\bm{w}}(s_i',a')$ and $||\bm{x}||_{\bm{A}} \triangleq \bm{x}^T\bm{A}^{-1}\bm{x}$ for $\bm{A}$ positive definite matrix.

@@

We assume a linear model for the $Q$-functions: $Q_{\bm{w}}(s,a) = \bm{w}^T\bm{\phi}(s,a)$. Here $\bm{\phi}$ is a $K$-dimensional feature vector. Then, the solution to the optimization problem of Eq. \eqref{eq:optim} can be computed in closed form as follows:
\begin{equation}
\bm{w}^* = \left( \bm{A}^T\bm{A} + \bm{\Sigma}^{-1} \right)^{-1} \left( \bm{A}^T\bm{b} + \bm{\Sigma}^{-1}\bm{\mu} \right) 
\end{equation}
where $\bm{A}$ is an $N\times K$ matrix containing the feature vectors at each data point $(s_i,a_i)$ and $\bm{b}$ is an $N$-dimensional vector containing their targets $y_i$.

>>>>

We define the Bellman residual of a $Q$-function $Q$ as $B(Q) \triangleq TQ - Q$. Notice that a $Q$-function $Q$ is optimal if, and only if, $B(Q)(s,a) = 0$ for all $s,a$.

@@

Consider solving the target task given a dataset $D$ of $N$ samples. The posterior distribution over optimal $Q$-functions is:
\begin{equation}
P(Q \mid D) \propto P(D \mid Q)P(Q),
\end{equation}
where $P(D \mid Q)$ is the likelihood of observing the dataset $D$ given that $Q$ is optimal and $P(Q)$ is the prior distribution induced by $\mathcal{D}$. Computing the likelihood requires knowing the transition and reward models of the current task (i.e., the MDP model), which are not available in practice. However, since we are conditioned on the fact that $Q$ is optimal, we can derive a simple approximation. We consider the empirical Bellman error $||B(Q)||_{p,D}^p$ under the $l_p$-norm defined as:
\begin{equation}
||B(Q)||_{p,D}^p = \frac{1}{N}\sum_{i=1}^N \left\lvert r_i + \gamma\max_{a'}Q(s_i',a') - Q(s_i,a_i) \right\rvert^p
\end{equation}
Assuming states and actions to be drawn from a fixed joint distribution $\mu$, we have $||B(Q)||_{p,\mu}^p = 0$ whenever $Q$ is optimal. Then, the probability of observing a dataset $D$ clearly decreases exponentially as $||B(Q)||_{p,D}^p$, for  an optimal $Q$, increases. This can be seen, for instance, by applying Hoeffding's inequality. Thus, a natural way to model the likelihood $P(D \mid Q)$ is:
\begin{equation}
P(D \mid Q) \propto e^{-\lambda N ||B(Q)||_{p,D}^p}
\end{equation}
where $\lambda$ is a constant hyperparameter. Intuitively, this indicates that $D$ is more likely when it induces low empirical Bellman error under an optimal $Q$-function. Furthermore, as the number of available samples $N$ increases, the distribution becomes more peaked at zero since the empirical Bellman error converges to the true Bellman error. In the limit $N \rightarrow \infty$:
\begin{equation}
P(D \mid Q) \propto I_{\left\{ ||B(Q)||_{p,D}^p = 0 \right\}}
\end{equation}

@@

Taking the maximum of the log-posterior, we obtain the following optimization problem:
\begin{equation}
\min_{Q\in\mathcal{Q}} ||B(Q)||_{p,D}^p - \log P(Q) 
\end{equation}
Let us specify a particular hypothesis space $\mathcal{Q}$. We consider $Q$-functions $Q_{\bm{w}}$ parameterized by the vector $\bm{w}$. Our optimization problem becomes:
\begin{equation}
\min_{\bm{w}} ||B(\bm{w})||_{p,D}^p - \log P(\bm{w}) 
\end{equation}
We model the prior distribution over the optimal parameters $\bm{w}$ as a Gaussian. That is, we assume:
\begin{equation}
P(\bm{w}) = \mathcal{N}(\bm{\mu},\bm{\Sigma})
\end{equation}
Then, our optimization, adopting the $l_2$-norm, becomes:
\begin{equation} \label{eq:optim}
\min_{\bm{w}} ||B(\bm{w})||_{D}^2 + ||\bm{w}-\bm{\mu}||_{\bm{\Sigma}} = \min_{\bm{w}} \frac{1}{N}\sum_{i=1}^N\left\lvert y_i - Q_{\bm{w}}(s_i,a_i) \right\rvert^2 + ||\bm{w}-\bm{\mu}||_{\bm{\Sigma}}
\end{equation}
where $y_i = r_i + \gamma\max_{a'}Q_{\bm{w}}(s_i',a')$ and $||\bm{x}||_{\bm{A}} \triangleq \bm{x}^T\bm{A}^{-1}\bm{x}$ for $\bm{A}$ positive definite matrix.

@@

We assume a linear model for the $Q$-functions: $Q_{\bm{w}}(s,a) = \bm{w}^T\bm{\phi}(s,a)$. Here $\bm{\phi}$ is a $K$-dimensional feature vector. Then, the solution to the optimization problem of Eq. \eqref{eq:optim} can be computed in closed form as follows:
\begin{equation}
\bm{w}^* = \left( \bm{A}^T\bm{A} + \bm{\Sigma}^{-1} \right)^{-1} \left( \bm{A}^T\bm{b} + \bm{\Sigma}^{-1}\bm{\mu} \right) 
\end{equation}
where $\bm{A}$ is an $N\times K$ matrix containing the feature vectors at each data point $(s_i,a_i)$ and $\bm{b}$ is an $N$-dimensional vector containing their targets $y_i$. The resulting algorithm is a regularized version of LSVI.

\paragraph{Neural network}

We model $Q_{\bm{w}}(s,a)$ as a neural network with parameters $\bm{w}$. Then, the gradient of the objective function of Eq. \eqref{eq:optim} can be easily computed by standard backpropagation. The resulting algorithm is a regularized version of NFQI.

\subsection{Variational Inference for Gaussian Distributions}

TODO
